### 2017-11-15 paper/sections/1_introduction.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear keyspace access patterns and hot
spots that offer significant opportunity for optimization. We use the data
management and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different domain and service.  Our exploration of this space results in a
dynamically sized cache policy that, for our initial conditions, sacrifices
negligible performance while using only 28\% of the memory required by our
hand-tuned cache.

@@

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a ParSplice cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}. The solid lines are
the existing cache management policies, exposed as configurations in ParSplice.
The default ParSplice configuration uses an unlimited sized cache, shown by the
``No Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where memory is precious and a common goal
is to keep memory for such data structures below 3\%\footnote{Empirically, we
find this threshold works well for most applications}.  While users can
configure ParSplice to resize the cache when it reaches a certain threshold,
this solution requires tuning and parameter sweeps to find the optimal cache
size that sacrifices negligible performance.  Furthermore, this threshold
changes with different initial configurations and clusters, as shown by our
keyspace analysis in Section~\S\ref{sec:parsplice-keyspace-analysis}.  The
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the expense of performance.  Our
dynamically sized cache, shown by the dashed line in
Figure~\ref{fig:cache-management}, detects keyspace access patterns and
re-sizes the cache accordingly. The design of this policy is driven by a
detailed analysis of the key-value accesses over the course of a long running
simulation across a variety of initial conditions.  Without tuning or parameter
sweeps, our solution saves more memory than a hand-tuned cache without
sacrificing performance.  More importantly, it works for a variety initial
conditions without changing the policy itself.

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable and facilitating rapid
  development. Changing the policy in use is critical in applications such as
  ParSplice that have alternating stable and chaotic keyspace access patterns
  over the course of a long-running simulation.

>>>>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot
spots that offer significant opportunity for optimization. We use the data
management and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different domain and service.  Our exploration of this space results in a
dynamically sized cache policy that, for our initial conditions, sacrifices
negligible performance while using only 28\% of the memory required by our
hand-tuned cache.

@@

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a ParSplice cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}. The solid lines are
the existing cache management policies, exposed as configurations in ParSplice.
The default ParSplice configuration uses an unlimited sized cache, shown by the
``No Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where memory is precious and a common goal
is to keep memory for such data structures below 3\%\footnote{Empirically, we
find this threshold works well for most applications}.  While users can
configure ParSplice to resize the cache when it reaches a certain threshold,
this solution requires tuning and parameter sweeps to find the optimal cache
size that sacrifices negligible performance.  Furthermore, this threshold
changes with different initial configurations and clusters, as shown by our
keyspace analysis in Section~\S\ref{sec:parsplice-keyspace-analysis}.  The
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the expense of performance.  Our
dynamically sized cache, shown by the dashed line in
Figure~\ref{fig:cache-management}, detects key access patterns and
re-sizes the cache accordingly. The design of this policy is driven by a
detailed analysis of the key-value accesses over the course of a long running
simulation across a variety of initial conditions.  Without tuning or parameter
sweeps, our solution saves more memory than a hand-tuned cache without
sacrificing performance.  More importantly, it works for a variety initial
conditions without changing the policy itself.

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable and facilitating rapid
  development. Changing the policy in use is critical in applications such as
  ParSplice that have alternating stable and chaotic key access patterns
  over the course of a long-running simulation.

### 2017-11-15 paper/sections/4_dynamic-policies.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Detecting these superbasins can lead to more effective cache management
strategies because the height of the keyspace accesses is ``how much" of cache
to evict and the width of the keyspace accesses is ``when" to evict values from
the cache.  The zoomed portion of Figure~\ref{fig:keyspace-zoomed} shows how a
single superbasin affects the key accesses. Moving along the \(x\) axis shows
that the number of unique keys accessed over time grows while moving along the
\(y\) axis shows that early keys are accessed more often.  Overall, superbasins
are never re-visited because the simulation only adds molecules; we can never
reach a state with less molecules. This is why keys are never re-accessed.
Despite these patterns, the following characteristics of superbasins make it
hard to detect them:

@@

  \item Statistics: decay on each key counts down until 0 0-valued keys are
evicted.  ``history-of-keyspace", set to 10 seconds, to evict keys.

@@

Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
patterns. Iterating backwards through the keyspace trace, at each time step we
find the lowest key ID and compare against the lowest key ID we have seen so
far (Line 7). We also maintain the top and bottom of each fan (Line 13) so we
can tell the ``how much" policy the number of keys to evict (Line 19).  The
algorithm iterates backwards over the keyspace trace because a change in the
minimum value signals a new fan. No signal exists iterating left to right, as
the maximum value always increases and the minimum values at the bottom of each
fan are sparse.  For example, the maximum distance between values along the
bottom edge of the zoomed fan in Figure~\ref{fig:keyspace-zoomed} is 125
seconds, while the maximum distance between minimum values for the fan before
is 0 seconds. As a result of this sparseness, iterating left to right requires
a ``window size" parameter to determine when we think a minimum value will not
show up again.  We exchange runtime, in the form of an O(\(n\)) algorithm where
\(n\) is the number events, for simplicity because this approach avoids adding
new thresholds for fan detection ({\it e.g.}, space between key accesses, space
between key IDs, and window size of consecutive key accesses).

>>>>

Detecting these superbasins can lead to more effective cache management
strategies because the height of the groups of key accesses is ``how much" of cache
to evict and the width of the groups of key accesses is ``when" to evict values from
the cache.  The zoomed portion of Figure~\ref{fig:keyspace-zoomed} shows how a
single superbasin affects the key accesses. Moving along the \(x\) axis shows
that the number of unique keys accessed over time grows while moving along the
\(y\) axis shows that early keys are accessed more often.  Overall, superbasins
are never re-visited because the simulation only adds molecules; we can never
reach a state with less molecules. This is why keys are never re-accessed.
Despite these patterns, the following characteristics of superbasins make it
hard to detect them:

@@

  \item Statistics: decay on each key counts down until 0 0-valued keys are
evicted.  ``history-of-key-accesses", set to 10 seconds, to evict keys.

@@

Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
patterns. Iterating backwards through the key access trace, at each time step we
find the lowest key ID and compare against the lowest key ID we have seen so
far (Line 7). We also maintain the top and bottom of each fan (Line 13) so we
can tell the ``how much" policy the number of keys to evict (Line 19).  The
algorithm iterates backwards over the key access trace because a change in the
minimum value signals a new fan. No signal exists iterating left to right, as
the maximum value always increases and the minimum values at the bottom of each
fan are sparse.  For example, the maximum distance between values along the
bottom edge of the zoomed fan in Figure~\ref{fig:keyspace-zoomed} is 125
seconds, while the maximum distance between minimum values for the fan before
is 0 seconds. As a result of this sparseness, iterating left to right requires
a ``window size" parameter to determine when we think a minimum value will not
show up again.  We exchange runtime, in the form of an O(\(n\)) algorithm where
\(n\) is the number events, for simplicity because this approach avoids adding
new thresholds for fan detection ({\it e.g.}, space between key accesses, space
between key IDs, and window size of consecutive key accesses).

### 2017-11-15 paper/sections/5_file-systems.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% Load balancing in FSs
Mantle was designed for file system metadata load balancing across a cluster of
dedicated metadata servers, where spreading requests across servers improves
performance. But another technique to reduce request load is caching. If
clients and servers maintain a consistent cache, the client can do operations
locally without contacting the metadata server.  The caches in CephFS have the
same performance and utilization trade-off as ParSplice, where large caches
improve performance at the expense of a higher memory footprint. High memory
utilization is problematic when a single metadata server maintains consistent
caches with many clients.  Using cache management policies from the previous
two sections has two benefits for load balancing: (1) increases the capacity of
a single metadata server, and (2) helps identify which parts of the cache to
keep local to a server. For example, applying the keyspace access pattern
detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
system access pattern has the potential to identify when to migrate keys and
how many keys to migrate.

@@

% other use cases caches/tiers 
Ceph has many other data management techniques that would benefit from the
caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
automatically size and manage cache
tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain keyspace access
traces. We hypothesize that since this is all software defined caching,
something more clever than LRU would improve cache utilization without
sacrificing too much performance.

>>>>

% Load balancing in FSs
Mantle was designed for file system metadata load balancing across a cluster of
dedicated metadata servers, where spreading requests across servers improves
performance. But another technique to reduce request load is caching. If
clients and servers maintain a consistent cache, the client can do operations
locally without contacting the metadata server.  The caches in CephFS have the
same performance and utilization trade-off as ParSplice, where large caches
improve performance at the expense of a higher memory footprint. High memory
utilization is problematic when a single metadata server maintains consistent
caches with many clients.  Using cache management policies from the previous
two sections has two benefits for load balancing: (1) increases the capacity of
a single metadata server, and (2) helps identify which parts of the cache to
keep local to a server. For example, applying the key access pattern
detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
system access pattern has the potential to identify when to migrate keys and
how many keys to migrate.

@@

% other use cases caches/tiers 
Ceph has many other data management techniques that would benefit from the
caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
automatically size and manage cache
tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain key access
traces. We hypothesize that since this is all software defined caching,
something more clever than LRU would improve cache utilization without
sacrificing too much performance.
